Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-11 doc/manuscript/ESSDmanuscript.Rmd
planning/organizing (+ misc changes)
@mawilliams99 , FYI. Changes include the following:
- inserted upfront r chunk to define table & figure numbers, insert relevant statistics (to be automated)
-inserted placeholder figures and tables
- added/ edited some content

   CURRENT TABLE 2 GOES HERE

@@

@@

@@

To improve the data accessible for C accounting, the IPCC created the Emission Factor Database (EFDB; [https://www.ipcc-nggip.iges.or.jp/EFDB/main.php](https://www.ipcc-nggip.iges.or.jp/EFDB/main.php)), which is intended as a recognized library of emission factors and other parameters that can be used for estimating greenhouse gas emissions and removals.
The EFDB can be used both for efforts to tally a nation's intended or accomplished greenhouse gas reductions, or as a basis of comparison for external parties to evaluate these inventories. 
The EFDB encourages researchers to submit estimates of emission factors or other related parameters from peer-reviewed journal papers or other accepted sources for inclusion in the database. 
In the case of forests, emission factors include carbon stocks and increments within various pools (Table&nbsp;\@ref(table_variables)).

@@

**Here, we**
(1) review IPCC definitions of relevant carbon stocks and increments;
(2) describe mapping of ForC to IPCC's EFDB;
(3) describe updates to ForC  (ForC v4.0), most of which were implemented to facilitate data transfer to EFDB;
(4) summarize the data in ForC that's relevant to EFDB and records that have been transferred to date; and
(5) provide recommendations for improving data collection, analysis, database, and accounting.

@@

## Carbon stocks
Forest ecosystem C stocks may be parsed into pools in various ways. 
IPCC parses into biomass (aboveground and belowground), dead organic matter (dead wood and litter), and soil organic matter (Table&nbsp;\@ref(table_variables)).
Quantifying these requires a one-time measurement.

```{r table_variables, eval = TRUE, echo=FALSE, warning=FALSE}
C_pools_table <- read.csv("figures_tables/C_pools.csv")
names(C_pools_table) <- stringr::str_replace_all(names(C_pools_table), "\\.", " ")

library(knitr)
library(kableExtra)
knitr::kable(C_pools_table, format = "latex",
             longtable = FALSE,
             caption = "\\textbf{Variables with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  row_spec(0, bold = TRUE)
```

@@

\clearpage
```{r, out.width = "15cm", echo = FALSE, fig.cap = "\\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; ForC variables corresponding to the stock, increment, influx and outflux; and relationships among them.} In many cases, the match of ForC variables to IPCC criteria depends upon measurement protocols (e.g., minimum DBH). Additional caveats are as follows: 1- assumes that change in foliage biomass is negligible (see note 7); 2- incomplete: excludes large branch fall; also, under IPCC definitions, outflux from aboveground biomass should include all sizes, influx to deadwood should include only above the minimum diameter chosen for dead wood; 3- incomplete: excludes belowground components;  4-incomplete: excludes breakage into pieces less than dead wood threshold size; 5-incomplete: excludes woody mortality of stems <10 cm DBH, decomposition of dead wood (aboveground and coarse roots) into sizes classified as litter, may exclude branch fall; 6- measurements often limited to decomposition of relatively fine litter and may exclude branches and stems below the dead wood size threshold and/or the more decomposed layers of the O horizon; 7 - foliage production is generally measured by collecting leaf-fall, a method that assumes that the influx = outflux (foliage biomass is roughly constant year-to-year); 8 - excludes branch fall, which is necessary for a full accounting of woody productivity but is typically assumed negligible for calculations of net biomass change."}
#this is for a two-column figure. for one-column figure, use {r, out.width = "8.3cm", echo = FALSE, fig.cap = "one column figure"}
knitr::include_graphics("figures_tables/C_variable_mapping.png")
```

@@

Mapping of variables is shown in Fig. 1

@@

@@

**figure: map of relevant ForC data with underlying FAO ecozones**

**(summarize the data in ForC that's relevant to EFDB, identifying gaps)**

dead wood and litter comparisons will be particularly interesting, as IPCC values are based on just a handful of references for each climate zone (table 2.2 in 2019 guidelines)

@@

**(Paragraph highlighting important gaps in variables / regions)**

Several variables of value to IPCC, including standing dead wood, woody mortality, delta.agb, are not calculated and presented as frequently as are AGB and ANPP_woody, even though they can readily be derived from the same census data. 
We recommend that researchers calculate and report these, as specified below.
Furthermore, there is an opportunity to fill data gaps by calculating these from existing census data.
For example, the core census protocol of the Forest Global Earth Observatory [ForestGEO; REFS] collects the data required to calculate standing dead wood, woody mortality, and delta.agb, but these have not been calculated and reported for all sites for which the appropriate number of censuses are available (n=1 for standing dead wood, n=2 for woody mortality and delta.agb) [but see REFS].

@@

## IPCC

An important challenge is that forests are changing rapidly, and data collected a decaade ago may no longer be relevant, particularly in the cases of C increments and fluxes.

>>>>

@@

   CURRENT TABLE 4 GOES HERE

@@

```{r eval = TRUE, echo=FALSE, warning=FALSE, results='hide'}
#Tables
table_pools =1
table_variables =2

#Figures
fig_variable_mapping=1
fig_map=2
fig_dominant_veg=3

#Statistics
n_records_ForC= NA # total n records in ForC
n_independent_records_ForC= NA #suspected duplicates removed
n_records_EFDB_variables= NA #n independent records relevant to EFDB
n_records_sent_to_EFDB = NA # n records sent to EFDB
n_records_posted_EFDB = NA # n records accepted and posted in EFDB
```

@@

To improve the data accessible for C accounting, the IPCC created the Emission Factor Database (EFDB; [https://www.ipcc-nggip.iges.or.jp/EFDB/main.php](https://www.ipcc-nggip.iges.or.jp/EFDB/main.php)), which is intended as a recognized library of emission factors and other parameters that can be used for estimating greenhouse gas emissions and removals.
The EFDB can be used both for efforts to tally a nation's intended or accomplished greenhouse gas reductions, or as a basis of comparison for external parties to evaluate these inventories. 
The EFDB encourages researchers to submit estimates of emission factors or other related parameters from peer-reviewed journal papers or other accepted sources for inclusion in the database. 
In the case of forests, emission factors include carbon stocks and increments within various pools [@ipcc_2019_2019].

@@

Here, we
(1) review IPCC definitions of relevant carbon stocks and increments;
(2) describe mapping of ForC to IPCC's EFDB;
(3) describe updates to ForC  (ForC v4.0), most of which were implemented to facilitate data transfer to EFDB;
(4) summarize the data in ForC that's relevant to EFDB and records that have been transferred to date; and
(5) provide recommendations for improving data collection, analysis, database, and accounting.

@@

## Carbon stocks
Forest ecosystem C stocks may be parsed into pools in various ways. 
IPCC parses into biomass (aboveground and belowground), dead organic matter (dead wood and litter), and soil organic matter (Table `r table_pools`).
Quantifying these requires a one-time measurement.

```{r table_pools, eval = TRUE, echo=FALSE, warning=FALSE}
C_pools_table <- read.csv("figures_tables/C_pools.csv")
names(C_pools_table) <- stringr::str_replace_all(names(C_pools_table), "\\.", " ")

library(knitr)
library(kableExtra)
knitr::kable(C_pools_table, format = "latex",
             longtable = FALSE,
             caption = "\\textbf{IPCC-defined forest carbon pools with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  row_spec(0, bold = TRUE)
```

@@

\clearpage
```{r fig_variable_mapping, out.width = "15cm", echo = FALSE, fig.cap = "\\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; ForC variables corresponding to the stock, increment, influx and outflux; and relationships among them.} In many cases, the match of ForC variables to IPCC criteria depends upon measurement protocols (e.g., minimum DBH). Additional caveats are as follows: 1- assumes that change in foliage biomass is negligible (see note 7); 2- incomplete: excludes large branch fall; also, under IPCC definitions, outflux from aboveground biomass should include all sizes, influx to deadwood should include only above the minimum diameter chosen for dead wood; 3- incomplete: excludes belowground components;  4-incomplete: excludes breakage into pieces less than dead wood threshold size; 5-incomplete: excludes woody mortality of stems <10 cm DBH, decomposition of dead wood (aboveground and coarse roots) into sizes classified as litter, may exclude branch fall; 6- measurements often limited to decomposition of relatively fine litter and may exclude branches and stems below the dead wood size threshold and/or the more decomposed layers of the O horizon; 7 - foliage production is generally measured by collecting leaf-fall, a method that assumes that the influx = outflux (foliage biomass is roughly constant year-to-year); 8 - excludes branch fall, which is necessary for a full accounting of woody productivity but is typically assumed negligible for calculations of net biomass change."}
#this is for a two-column figure. for one-column figure, use {r, out.width = "8.3cm", echo = FALSE, fig.cap = "one column figure"}
knitr::include_graphics("figures_tables/C_variable_mapping.png")
```

@@

Mapping of variables is shown in Fig. `r `fig_variable_mapping`.

@@

Because ForC v4.0 contained known duplicate records, we used R scripts to remove likely duplicates, as detailed in @anderson-teixeira_carbon_2021.
Henceforth, we refer to the records with duplicates removed as "independent records".

@@

## ForC v4.0 contents

As of `r format(Sys.time(), '%B %d, %Y')`, ForC (v4.0) contained `r n_independent_records_ForC` independent records (`r n_records_ForC` total), `r n_records_EFDB_variables` of which were for variables relevant to EFDB (Fig. `r fig_variable_mapping`).
These records were distributed across all forested continents and ecozones, with particularly high concentrations in *[ecozone(s)]* and low concentrations in *[ecozone(s)]* (Fig. `r fig_map`).
The most widely represented forest type was *[type]*, followed by *[type]* and *[type]* (Fig. `r fig_dominant_veg`).

```{r fig_map, out.width = "15cm", echo = FALSE, fig.cap = "\\textbf{Map of sites in ForC shaded by number of records relavent to (circles) and transferred to (triangles) EFDB.} PLACEHOLDER. We need to add records transferred to EFDB. We also need to limit map to variables relavent to EFDB. It would also be good to add an underlying map of FAP ecozones"}
#this is for a two-column figure. for one-column figure, use {r, out.width = "8.3cm", echo = FALSE, fig.cap = "one column figure"}
knitr::include_graphics("figures_tables/PLACEHOLDER_World_Map_of_sites.png")
```

```{r fig_dominant_veg, out.width = "15cm", echo = FALSE, fig.cap = "\\textbf{Historgram of number of records in ForC relavent to (grey) and transferred to (black) EFDB.} PLACEHOLDER. We need to add axis labels and the two classes of records. "}
#this is for a two-column figure. for one-column figure, use {r, out.width = "8.3cm", echo = FALSE, fig.cap = "one column figure"}
knitr::include_graphics("figures_tables/PLACEHOLDER_Williams_Histogram_of_dominant_vegetation.png")
```

ForC contained data for **##** of the **##** variables relevant to EFDB (Table `r table_variables`, Fig. `r fig_variable_mapping`).

```{r table_variables, eval = TRUE, echo=FALSE, warning=FALSE}
C_variables_table <- read.csv("figures_tables/C_variables.csv")
names(C_variables_table) <- stringr::str_replace_all(names(C_variables_table), "\\.", " ")

library(knitr)
library(kableExtra)
knitr::kable(C_variables_table, format = "latex",
             longtable = FALSE,
             caption = "\\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. ") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  row_spec(0, bold = TRUE)
```

## Data transfers to EFDB

As of `r format(Sys.time(), '%B %d, %Y')`<!--inserts date upon knit-->, we had reviewed and sent `r n_records_sent_to_EFDB`, `r n_records_posted_EFDB` of which have been reviewed, accepted, and posted (Fig. `r fig_map`-`r fig_dominant_veg`, `Table r table_variables`).
*[DETAILS]*

@@

*(Paragraph highlighting important gaps in variables / regions)*

<!--dead wood and litter comparisons will be particularly interesting, as IPCC values are based on just a handful of references for each climate zone (table 2.2 in 2019 guidelines)-->

Several variables of value to IPCC, including standing dead wood, woody mortality, delta.agb, are not calculated and presented as frequently as are AGB and ANPP_woody, even though they can readily be derived from the same census data. 
We recommend that researchers calculate and report these, following the reporting guidelines specified below.
Furthermore, there is an opportunity to fill data gaps by calculating these from existing census data.
For example, the core census protocol of the Forest Global Earth Observatory [ForestGEO; @anderson-teixeira_ctfsforestgeo_2015,@davies_forestgeo_2021] collects the data required to calculate standing dead wood, woody mortality, and delta.agb, but these have not been calculated and reported for all sites for which the appropriate number of censuses are available <!--"(n=1 for standing dead wood, n=2 for woody mortality and delta.agb)" -- previous text commented out because I think it's out of date-->[but see @piponiot_distribution_2022, @refs<!--insert other refs including these variables-->].

@@

## IPCC protocol considerations

An important challenge is that forests are changing rapidly, and data collected a decade ago may no longer be relevant, particularly in the cases of C increments and fluxes.

### 2022-04-11 doc/manuscript/ESSDmanuscript.tex
planning/organizing (+ misc changes)
@mawilliams99 , FYI. Changes include the following:
- inserted upfront r chunk to define table & figure numbers, insert relevant statistics (to be automated)
-inserted placeholder figures and tables
- added/ edited some content

To improve the data accessible for C accounting, the IPCC created the
Emission Factor Database (EFDB;
\url{https://www.ipcc-nggip.iges.or.jp/EFDB/main.php}), which is
intended as a recognized library of emission factors and other
parameters that can be used for estimating greenhouse gas emissions and
removals. The EFDB can be used both for efforts to tally a nation's
intended or accomplished greenhouse gas reductions, or as a basis of
comparison for external parties to evaluate these inventories. The EFDB
encourages researchers to submit estimates of emission factors or other
related parameters from peer-reviewed journal papers or other accepted
sources for inclusion in the database. In the case of forests, emission
factors include carbon stocks and increments within various pools
(Table~@ref(table\_variables)).

@@

\textbf{Here, we} (1) review IPCC definitions of relevant carbon stocks
and increments; (2) describe mapping of ForC to IPCC's EFDB; (3)
describe updates to ForC (ForC v4.0), most of which were implemented to
facilitate data transfer to EFDB; (4) summarize the data in ForC that's
relevant to EFDB and records that have been transferred to date; and (5)
provide recommendations for improving data collection, analysis,
database, and accounting.

@@

Forest ecosystem C stocks may be parsed into pools in various ways. IPCC
parses into biomass (aboveground and belowground), dead organic matter
(dead wood and litter), and soil organic matter
(Table~@ref(table\_variables)). Quantifying these requires a one-time
measurement.

@@

\caption{\label{tab:table_variables}\textbf{Variables with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{pool} & \textbf{definition} & \textbf{major sources of estimate variation} & \textbf{IPCC guidance}\\
\hline
aboveground biomass & all biomass of living vegetation, both woody and herbaceous, above the soil & allometry, min dbh & acceptable to exclude understory\\
\hline
belowground biomass & all biomass of live roots & allometry, min dbh, assumed ratio of belowground to aboveground biomass (IPCC table 4.4) & fine roots may be excluded when they cannot be distinguished empirically from soil organic matter or litter\\
\hline
dead wood & all non-living woody biomass not contained in the litter, either standing, lying on the ground, or in the soil & min dbh,  ... & default min dbh = 10cm, but may be chosen by country\\
\hline
litter & all non-living biomass with a size greater than the limit for soil organic matter  and less than the minimum diameter chosen for dead wood, lying dead, in various states of decomposition above or within the mineral or organic soil & min dbh for dead wood , .. & includes entire O horizon\\
\hline
soil organic matter & organic carbon in mineral soils to a specified depth & sampling depth & default sampling depth = 30cm, but may be chosen by country\\
\hline
\end{tabu}
\end{table}

@@

\clearpage
\begin{figure}
\includegraphics[width=15cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; ForC variables corresponding to the stock, increment, influx and outflux; and relationships among them.} In many cases, the match of ForC variables to IPCC criteria depends upon measurement protocols (e.g., minimum DBH). Additional caveats are as follows: 1- assumes that change in foliage biomass is negligible (see note 7); 2- incomplete: excludes large branch fall; also, under IPCC definitions, outflux from aboveground biomass should include all sizes, influx to deadwood should include only above the minimum diameter chosen for dead wood; 3- incomplete: excludes belowground components;  4-incomplete: excludes breakage into pieces less than dead wood threshold size; 5-incomplete: excludes woody mortality of stems <10 cm DBH, decomposition of dead wood (aboveground and coarse roots) into sizes classified as litter, may exclude branch fall; 6- measurements often limited to decomposition of relatively fine litter and may exclude branches and stems below the dead wood size threshold and/or the more decomposed layers of the O horizon; 7 - foliage production is generally measured by collecting leaf-fall, a method that assumes that the influx = outflux (foliage biomass is roughly constant year-to-year); 8 - excludes branch fall, which is necessary for a full accounting of woody productivity but is typically assumed negligible for calculations of net biomass change.}\label{fig:unnamed-chunk-1}
\end{figure}

@@

Mapping of variables is shown in Fig. 1

@@

@@

\textbf{figure: map of relevant ForC data with underlying FAO ecozones}

\textbf{(summarize the data in ForC that's relevant to EFDB, identifying
gaps)}

dead wood and litter comparisons will be particularly interesting, as
IPCC values are based on just a handful of references for each climate
zone (table 2.2 in 2019 guidelines)

@@

\textbf{(Paragraph highlighting important gaps in variables / regions)}

Several variables of value to IPCC, including standing dead wood, woody
mortality, delta.agb, are not calculated and presented as frequently as
are AGB and ANPP\_woody, even though they can readily be derived from
the same census data. We recommend that researchers calculate and report
these, as specified below. Furthermore, there is an opportunity to fill
data gaps by calculating these from existing census data. For example,
the core census protocol of the Forest Global Earth Observatory
{[}ForestGEO; REFS{]} collects the data required to calculate standing
dead wood, woody mortality, and delta.agb, but these have not been
calculated and reported for all sites for which the appropriate number
of censuses are available (n=1 for standing dead wood, n=2 for woody
mortality and delta.agb) {[}but see REFS{]}.

@@

\subsection{IPCC}

An important challenge is that forests are changing rapidly, and data
collected a decaade ago may no longer be relevant, particularly in the
cases of C increments and fluxes.

@@

CURRENT TABLE 2 GOES HERE

@@

CURRENT TABLE 3 GOES HERE
\noappendix

>>>>

To improve the data accessible for C accounting, the IPCC created the
Emission Factor Database (EFDB;
\url{https://www.ipcc-nggip.iges.or.jp/EFDB/main.php}), which is
intended as a recognized library of emission factors and other
parameters that can be used for estimating greenhouse gas emissions and
removals. The EFDB can be used both for efforts to tally a nation's
intended or accomplished greenhouse gas reductions, or as a basis of
comparison for external parties to evaluate these inventories. The EFDB
encourages researchers to submit estimates of emission factors or other
related parameters from peer-reviewed journal papers or other accepted
sources for inclusion in the database. In the case of forests, emission
factors include carbon stocks and increments within various pools
\citep{ipcc_2019_2019}.

@@

Here, we (1) review IPCC definitions of relevant carbon stocks and
increments; (2) describe mapping of ForC to IPCC's EFDB; (3) describe
updates to ForC (ForC v4.0), most of which were implemented to
facilitate data transfer to EFDB; (4) summarize the data in ForC that's
relevant to EFDB and records that have been transferred to date; and (5)
provide recommendations for improving data collection, analysis,
database, and accounting.

@@

Forest ecosystem C stocks may be parsed into pools in various ways. IPCC
parses into biomass (aboveground and belowground), dead organic matter
(dead wood and litter), and soil organic matter (Table 1). Quantifying
these requires a one-time measurement.

@@

\caption{\label{tab:table_pools}\textbf{IPCC-defined forest carbon pools with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{pool} & \textbf{definition} & \textbf{major sources of estimate variation} & \textbf{IPCC guidance}\\
\hline
aboveground biomass & all biomass of living vegetation, both woody and herbaceous, above the soil & allometry, min dbh & acceptable to exclude understory\\
\hline
belowground biomass & all biomass of live roots & allometry, min dbh, assumed ratio of belowground to aboveground biomass (IPCC table 4.4) & fine roots may be excluded when they cannot be distinguished empirically from soil organic matter or litter\\
\hline
dead wood & all non-living woody biomass not contained in the litter, either standing, lying on the ground, or in the soil & min dbh,  ... & default min dbh = 10cm, but may be chosen by country\\
\hline
litter & all non-living biomass with a size greater than the limit for soil organic matter  and less than the minimum diameter chosen for dead wood, lying dead, in various states of decomposition above or within the mineral or organic soil & min dbh for dead wood , .. & includes entire O horizon\\
\hline
soil organic matter & organic carbon in mineral soils to a specified depth & sampling depth & default sampling depth = 30cm, but may be chosen by country\\
\hline
\end{tabu}
\end{table}

@@

\clearpage
\begin{figure}
\includegraphics[width=15cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools quantified under IPCC accounting; ForC variables corresponding to the stock, increment, influx and outflux; and relationships among them.} In many cases, the match of ForC variables to IPCC criteria depends upon measurement protocols (e.g., minimum DBH). Additional caveats are as follows: 1- assumes that change in foliage biomass is negligible (see note 7); 2- incomplete: excludes large branch fall; also, under IPCC definitions, outflux from aboveground biomass should include all sizes, influx to deadwood should include only above the minimum diameter chosen for dead wood; 3- incomplete: excludes belowground components;  4-incomplete: excludes breakage into pieces less than dead wood threshold size; 5-incomplete: excludes woody mortality of stems <10 cm DBH, decomposition of dead wood (aboveground and coarse roots) into sizes classified as litter, may exclude branch fall; 6- measurements often limited to decomposition of relatively fine litter and may exclude branches and stems below the dead wood size threshold and/or the more decomposed layers of the O horizon; 7 - foliage production is generally measured by collecting leaf-fall, a method that assumes that the influx = outflux (foliage biomass is roughly constant year-to-year); 8 - excludes branch fall, which is necessary for a full accounting of woody productivity but is typically assumed negligible for calculations of net biomass change.}\label{fig:fig_variable_mapping}
\end{figure}

@@

Mapping of variables is shown in Fig. \texttt{r}fig\_variable\_mapping`.

@@

Because ForC v4.0 contained known duplicate records, we used R scripts
to remove likely duplicates, as detailed in
\citet{anderson-teixeira_carbon_2021}. Henceforth, we refer to the
records with duplicates removed as ``independent records''.

@@

\subsection{ForC v4.0 contents}

As of April 11, 2022, ForC (v4.0) contained NA independent records (NA
total), NA of which were for variables relevant to EFDB (Fig. 1). These
records were distributed across all forested continents and ecozones,
with particularly high concentrations in \emph{{[}ecozone(s){]}} and low
concentrations in \emph{{[}ecozone(s){]}} (Fig. 2). The most widely
represented forest type was \emph{{[}type{]}}, followed by
\emph{{[}type{]}} and \emph{{[}type{]}} (Fig. 3).

\begin{figure}
\includegraphics[width=15cm]{figures_tables/PLACEHOLDER_World_Map_of_sites} \caption{\textbf{Map of sites in ForC shaded by number of records relavent to (circles) and transferred to (triangles) EFDB.} PLACEHOLDER. We need to add records transferred to EFDB. We also need to limit map to variables relavent to EFDB. It would also be good to add an underlying map of FAP ecozones}\label{fig:fig_map}
\end{figure}

\begin{figure}
\includegraphics[width=15cm]{figures_tables/PLACEHOLDER_Williams_Histogram_of_dominant_vegetation} \caption{\textbf{Historgram of number of records in ForC relavent to (grey) and transferred to (black) EFDB.} PLACEHOLDER. We need to add axis labels and the two classes of records. }\label{fig:fig_dominant_veg}
\end{figure}

ForC contained data for \textbf{\#\#} of the \textbf{\#\#} variables
relevant to EFDB (Table 2, Fig. 1).

\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. }
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB} & \textbf{n posted to EFDB}\\
\hline
STOCKS & NA & NA & NA & NA & NA\\
\hline
biomass & NA & NA & NA & NA & NA\\
\hline
biomass\_ag & NA & NA & NA & NA & NA\\
\hline
biomass\_ag\_foliage & NA & NA & NA & NA & \vphantom{1} NA\\
\hline
biomass\_ag\_woody & NA & NA & NA & NA & NA\\
\hline
biomass\_root & NA & NA & NA & NA & NA\\
\hline
biomass\_root\_fine & NA & NA & NA & NA & NA\\
\hline
biomass\_root\_coarse & NA & NA & NA & NA & NA\\
\hline
deadwood & NA & NA & NA & NA & NA\\
\hline
deadwood\_standing & NA & NA & NA & NA & NA\\
\hline
deadwood\_down & NA & NA & NA & NA & NA\\
\hline
O.horizon & NA & NA & NA & NA & NA\\
\hline
SOM / SOC & NA & NA & NA & NA & NA\\
\hline
INCREMENTS & NA & NA & NA & NA & NA\\
\hline
delta.biomass & NA & NA & NA & NA & NA\\
\hline
delta.agb & NA & NA & NA & NA & NA\\
\hline
biomass\_ag\_foliage & NA & NA & NA & NA & NA\\
\hline
delta.biomass\_root & NA & NA & NA & NA & NA\\
\hline
delta.deadwood & NA & NA & NA & NA & NA\\
\hline
delta.deadwood\_standing & NA & NA & NA & NA & NA\\
\hline
delta.deadwood\_down & NA & NA & NA & NA & NA\\
\hline
delta.O.horizon & NA & NA & NA & NA & NA\\
\hline
delta.SOM / delta.SOC & NA & NA & NA & NA & NA\\
\hline
FLUXES & NA & NA & NA & NA & NA\\
\hline
NPP\_woody & NA & NA & NA & NA & NA\\
\hline
ANPP\_woody\_stem & NA & NA & NA & NA & NA\\
\hline
ANPP\_foliage & NA & NA & NA & NA & NA\\
\hline
BNPP\_root & NA & NA & NA & NA & NA\\
\hline
BNPP\_root\_fine & NA & NA & NA & NA & NA\\
\hline
BNPP\_root\_coarse & NA & NA & NA & NA & NA\\
\hline
woody\_mortality & NA & NA & NA & NA & NA\\
\hline
NPP\_litterfall & NA & NA & NA & NA & NA\\
\hline
R\_het\_litter & NA & NA & NA & NA & NA\\
\hline
R\_het\_soil & NA & NA & NA & NA & NA\\
\hline
TOTAL & NA & NA & NA & NA & NA\\
\hline
\end{tabu}
\end{table}

\subsection{Data transfers to EFDB}

As of April 11, 2022, we had reviewed and sent NA, NA of which have been
reviewed, accepted, and posted (Fig. 2-3,
\texttt{Table\ r\ table\_variables}). \emph{{[}DETAILS{]}}

@@

\emph{(Paragraph highlighting important gaps in variables / regions)}

Several variables of value to IPCC, including standing dead wood, woody
mortality, delta.agb, are not calculated and presented as frequently as
are AGB and ANPP\_woody, even though they can readily be derived from
the same census data. We recommend that researchers calculate and report
these, following the reporting guidelines specified below. Furthermore,
there is an opportunity to fill data gaps by calculating these from
existing census data. For example, the core census protocol of the
Forest Global Earth Observatory {[}ForestGEO;
\citet{anderson-teixeira_ctfsforestgeo_2015},\citet{davies_forestgeo_2021}{]}
collects the data required to calculate standing dead wood, woody
mortality, and delta.agb, but these have not been calculated and
reported for all sites for which the appropriate number of censuses are
available \citep[but see][\citet{refs}]{piponiot_distribution_2022}.

@@

\subsection{IPCC protocol considerations}

An important challenge is that forests are changing rapidly, and data
collected a decade ago may no longer be relevant, particularly in the
cases of C increments and fluxes.

@@

CURRENT TABLE 3 GOES HERE

@@

CURRENT TABLE 4 GOES HERE
\noappendix
